### 2016-06-24 GPCE2016/sections/CaseStudy.tex
Marco improvements.
I can not remember the macro to add comments, used @@@

@@

>>>>

{\bf @@@@@@@@@@@@FIX HERE, add AddPC and LitPC @@@@@@@}

@@

Smalltalk and Dart offer method cascading~[??];
at the price of introducing additional syntax,
this avoid the need for fluent setters;
method cascading (as setters in general) 
intrinsically rely  on an imperative setting.
Our approach supports both fluent setters and (functional) 
fluent withers.

### 2016-06-24 GPCE2016/sections/Implementation.tex
revise. refer to appendix material in external link.
Marco improvements.
I can not remember the macro to add comments, used @@@

Our automatic type refinement is a useful and convenient feature, but
not transparent to the heirs of the annotated interface.  They need to
be aware of the annotation semantics and provide the right type while
refining methods. To support heir coherence, we would need
to give up automatic type refinement, which is an essential part of IB programming.
However, the reader should note that Java libraries almost always break heir
coherence during evolution and still claim backward compatibility (false in
theory but statistically true in practice). In practice, adding any method to any
non-final class of a Java library is enough to break heir
coherence.  We think return type refinement breaks heir coherence ``less" than normal library evolution, and
if no automatic type-refinements are needed, then \mixin can claim a
form of heir coherence.
Section~\ref{sec:translation} provide a formal definition for our
safety claims.

\subsection{Limitations}
Our prototype implementation has certain limitations:
\begin{itemize}
\item Lombok allows writing handlers for either javac or ejc(Eclipse's own compiler).
Our current implementation only realizes ejc version. The implementation for
  the \texttt{javac} version is still missing.
\item Simple generics is supported:
type parameters can be used, but generic
method typing %is neither formalized in this paper nor 
not explicitly checked by \mixin, but simply delegated to the Java compiler.
\item
Lombok offers only limited/experimental support for separate compilation, that is
to access information of code defined in different files.
Limited by Lombok, \mixin 
%In the same spirit, we have a mature \mixin annotation, which
%does not support separate compilation yet: it 
requires that all
  related interfaces have to appear in a single Java file.
Reusing the logic inside the experimental Lombok annotation \lstinline{@Delegate},
we also offer a less polished annotation supporting
separate compilation for files in the same package.
%  has the same behavior as \mixin. A limitation is
%  that users need to put references to the super types
% of the annotated type together with the annotation, for instance,
%  \lstinline{@Delegate(types = Point2D.class)}.
%   Another limitation is the methods generated by \lstinline{@Delegate} will
%  not be visible in the Outline window, but they can still be auto-completed in Eclipse.
\end{itemize}

>>>>

Our automatic type refinement is a useful and convenient feature, but
not transparent to the heirs of the annotated interface.  They need to
be aware of the annotation semantics and provide the right type while
refining methods. To support heir coherence, we would need
to give up automatic type refinement, which is an essential part of IB programming.
However, the reader should note that Java libraries almost always break heir
coherence during evolution and still claim backward compatibility (false in
theory but statistically true in practice). In practice, adding any method to any
non-final class of a Java library is enough to break heir
coherence.  We think return type refinement breaks heir coherence ``less" than normal library evolution, and
if no automatic type-refinements are needed, then \mixin can claim a
form of heir coherence.
%Section~\ref{sec:translation} 
We provide a formal definition for our safety claims here~\footnote{See url?? Section B}.

\subsection{Limitations}
Our prototype implementation has certain limitations:
\begin{itemize}
\item Lombok allows writing handlers for either javac or ejc(Eclipse's own compiler).
Our current implementation only realizes ejc version. The implementation for
  the \texttt{javac} version is still missing.
\item Simple generics is supported:
type parameters can be used, but generic
method typing %is neither formalized in this paper nor 
not explicitly checked by \mixin, but simply delegated to the Java compiler.
\item
{\bf @@@English need improvement here@@@@@@@}
Lombok offers only limited/experimental support for separate compilation, that is
to access information of code defined in different files.
Limited by Lombok, \mixin 
%In the same spirit, we have a mature \mixin annotation, which
%does not support separate compilation yet: it 
requires that all
  related interfaces have to appear in a single Java file.
Reusing the logic inside the experimental Lombok annotation \lstinline{@Delegate},
we also offer a less polished annotation supporting
separate compilation for files in the same package.
%  has the same behavior as \mixin. A limitation is
%  that users need to put references to the super types
% of the annotated type together with the annotation, for instance,
%  \lstinline{@Delegate(types = Point2D.class)}.
%   Another limitation is the methods generated by \lstinline{@Delegate} will
%  not be visible in the Outline window, but they can still be auto-completed in Eclipse.
\end{itemize}

### 2016-06-24 GPCE2016/sections/Introduction.tex
Marco improvements.
I can not remember the macro to add comments, used @@@

In the presence of subtyping, abstract state operations often require
special care, as their types need to be refined. Object interfaces
provide support for type-refinement and can automatically produce code
that deals with type-refinement adequately. In contrast,
 verbose explicit type-refinement
 is typically required in CB.
  We believe that such verbosity hindered and slowed
down the discovery of useful programming patterns involving
type-refinement. A recently discovered
solution~\cite{eptrivially} to the Expression
Problem~\cite{wadler98expression} in Java-like languages, shows how easy it is to solve
the problem using only type-refinement. However it took nearly 20
years since the formulation of the problem for that
solution to be presented in the literature. In IB, due to
its emphasis on type-refinement, that solution should have been much
more obvious.

>>>>

Object interfaces
provide support for automatic type-refinement.
In contrast, in CB 
special care and verbose explicit type-refinement is be required to
produce code that deals with subtyping adequately.

  We believe that such verbosity hindered and slowed
down the discovery of useful programming patterns involving
type-refinement. A recently discovered
solution~\cite{eptrivially} to the Expression
Problem~\cite{wadler98expression} in Java-like languages, shows how easy it is to solve
the problem using only type-refinement. However it took nearly 20
years since the formulation of the problem for that
solution to be presented in the literature. In IB, due to
its emphasis on type-refinement, that solution should have been much
more obvious.

### 2016-06-24 GPCE2016/sections/Overview.tex
revise. refer to appendix material in external link.
Marco improvements.
I can not remember the macro to add comments, used @@@

This section illustrates how our programming style, supported by 
\mixinAnn{}, enables powerful programming idioms based on multiple
inheritance and type refinements.  We propose a standard example:
\Q@Animal@s with a 2-dimensional \Q@Point2D@ representing their
\Q@location@.  Some kinds of animals are \Q@Horses@ and \Q@Bird@s.
Birds can \Q@fly@, thus their locations need to be 3-dimensional
\Q@Point3D@s (field type refinement). Finally, we model \Q@Pegasus@
(a well-known creature in Greek mythology) as a kind of
\Q@Animal@ with the skills of both \Q@Horse@s and \Q@Bird@s (multiple
inheritance). A simple class diagram illustrating the basic system is
given on the left side of Figure~\ref{fig:pegasus}.\footnote{Some
  research argues in favor of using subtyping for modeling taxonomies,
  other research argues against this practice, we do not wish to take
  sides in this argument, but to provide an engaging example.}

@@

\paragraph{\Q@Point2D@: simple immutable data with fields}
Here we will illustrate how points are modelled with interfaces. In IB
we do not talk about state directly. Instead, state is accessed and
manipulated using abstract methods.  The usual approach to model
points in Java is to use a class with fields for the coordinates.
In Classless Java interfaces are used instead:

@@

\noindent The encoding over Java is now inconvenient: creating a new point object is cumbersome, even
with anonymous classes:

@@

\noindent More precisely, a field or factory parameter is generated for every
abstract method that takes no parameters (except for methods with special
names). An example of using \Q@Point2D@ is:
\begin{lstlisting}
Point2D p = Point2D.of(42,myPoint.y());
\end{lstlisting}
\noindent where we return a new point, using \Q@42@ as x-coordinate,
and taking all the other information (only \Q@y@ in this case) from
another point.

@@

\begin{lstlisting}
interface Point2D { int x(); int y();
    static Point2D of(int _x, int _y){ return new Point2D(){
        int x=_x; int y=_y;
        public int x(){return x;}   public int y(){ return y; }
        Point2D withX(int val){ return Point2D.of(val,this.y()); }
        Point2D withY(int val){ return Point2D.of(this.x(),val); }
    }; }
  Point2D withX(int val);    Point2D withY(int val); }
\end{lstlisting}
\end{comment}
Moreover, if the programmer wants a different implementation, he may
provide an alternative implementation using \Q@default@ methods. For example:
\begin{lstlisting}
@Obj interface Point2D {
    int x(); int y();
    default Point2D withX(int val){ /*myCode*/ }
    default Point2D withY(int val){ /*myCode*/ } }
\end{lstlisting}

%\begin{comment}
%\marco{I re-enabled this code, I think is needed for understandability}
\noindent is expanded into
\begin{lstlisting}
interface Point2D {
    int x(); int y();
    default Point2D withX(int val){ /*myCode*/ }
    default Point2D withY(int val){ /*myCode*/ }
    static Point2D of(int _x, int _y){
      return new Point2D(){
        int x=_x;    int y=_y;
        public int x(){return x;}
        public int y(){return y;} }; } }
\end{lstlisting}

@@

>>>>

This section illustrates how our programming style, supported by 
\mixinAnn{}, enables powerful programming idioms based on multiple
inheritance and type refinements.  We propose a standard example:
\Q@Animal@s with a 2-dimensional \Q@Point2D@ representing their
\Q@location@, subtypes \Q@Horses@, \Q@Bird@s, and \Q@Pegasus@.
Birds can \Q@fly@, thus their locations need to be 3-dimensional
\Q@Point3D@s (field type refinement). We model \Q@Pegasus@
(a well-known creature in Greek mythology) as a kind of
\Q@Animal@ with the skills of both \Q@Horse@s and \Q@Bird@s (multiple
inheritance). A simple class diagram illustrating the basic system is
given on the left side of Figure~\ref{fig:pegasus}.\footnote{Some
  research argues in favor of using subtyping for modeling taxonomies,
  other research argues against this practice, we do not wish to take
  sides in this argument, but to provide an engaging example.}

@@

\paragraph{\Q@Point2D@: simple immutable data with fields}
Here we will illustrate how points are modelled with interfaces.
 In IB
 state is accessed and
manipulated using abstract methods.  The usual approach to model
points in Java is to use a class with fields for the coordinates.
In Classless Java interfaces are used instead:

@@

\noindent The encoding over pure Java is now inconvenient: creating a new point object is cumbersome, even
with anonymous classes:

@@

\noindent More precisely, a field or factory parameter is generated for every
abstract method that takes no parameters.
%(except for methods with special
%names).
 An example of using \Q@Point2D@ is:
\begin{lstlisting}
Point2D p = Point2D.of(42,myPoint.y());
\end{lstlisting}
\noindent where we return a new point, using \Q@42@ as x-coordinate,
and taking all the other information (only \Q@y@ in this case) from
another point.

@@

\begin{lstlisting}
interface Point2D { int x(); int y();
    static Point2D of(int _x, int _y){ return new Point2D(){
        int x=_x; int y=_y;
        public int x(){return x;}   public int y(){ return y; }
        Point2D withX(int val){ return Point2D.of(val,this.y()); }
        Point2D withY(int val){ return Point2D.of(this.x(),val); }
    }; }
  Point2D withX(int val);    Point2D withY(int val); }
\end{lstlisting}
\end{comment}
Moreover, if the programmer wants a different implementation, they may
provide an alternative implementation using \Q@default@ methods. For example:
\begin{lstlisting}
@Obj interface Point2D {
    int x(); int y();
    default Point2D withX(int val){ /*myCode*/ }
    Point2D withY(int val); }
\end{lstlisting}

%\begin{comment}
%\marco{I re-enabled this code, I think is needed for understandability}
\noindent is expanded into
\begin{lstlisting}
interface Point2D {
    int x(); int y();
    default Point2D withX(int val){ /*myCode*/ }
    Point2D withY(int val);
    static Point2D of(int _x, int _y){
      return new Point2D(){
        int x=_x;    int y=_y;
        public int x(){return x;}
        public int y(){return y;} 
        public Point2D withY(int val){
          return of(x(),val);}  }; } }
\end{lstlisting}

@@

\paragraph{Property updater as Injections}

By declaring fields in Newspeak and other languages,
the programmers specifies field signatures and (critically) the intention of storing such information. Then the language generate getter and setter.
Once state is abstracted away,
it is well known that is possible to replace state
access with computation, but the type of the field stays the same.

Our fluid approach to state is specular:
the intention of storing the information is not expressed by the programmer and set in stone, 
but can vary by inheritance:
We generate the idea of the field
just starting from getters and setters signatures.

In this case, we leverage on our fluid state to changing the underling type of the field,
 and our \Q@with@ is providing
the right injection from the old type to the new one.
